Waving and swaying

### 2014-03-03 Paper/grass.tex
update March3

transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc. 
The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami (Ackerman and Okubo, 1993), plays a central role
in the recruitment of microscopic marine organism such as blue mussel larvae. The hydrodynamic mechanism underlying monami is the focus of this paper. 

@@

Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami (Inoue 1955, Raupach 1996) have also been observed.

@@

Due to these considerable differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 

@@

Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy (Ghisalberti 2002, Raupach 1996) due to

@@

consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical 
Reynolds number (Grizzle 1996) for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires 

@@

In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grass

@@

instability associated with strong shear near the canopy top but also through shear instability of flow above grass canopies arising due to existence of flow boundary.
We further show that inclusion of drag due to grass predicts a threshold condition for waving where 
threshold is characterized by Reynolds number. We confirm our analysis by comparing prediction from theory with field and experimental observations.

@@

primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depends weakly on flexibility of grass

@@

Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices. We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 

@@

The steady uniform flow resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. steady state flow driven by   

@@

\small$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$\normalsize arising from balance of drag force with pressure gradient with in grass and a parabolic velocity 

@@

as shear layer in previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 

@@

with drag along with match of shear at grass tip exerted by flow in unvegetated region, which leads to shear layer 
width \small $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity.

@@

Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0 = \frac{dP/dxH^2}{\mu0}$.

@@

\scriptstyle{
Re\left[\del_t+\del_x U \right]\grad^2\psi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2D_{drag}}{R_e}U\psi_y\right)
}

@@

where $R_{e}= \frac{\rho U_0 H}{\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is 
non-dimensional drag. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, we get generalized orr-sommerfield equation 
which is an eigenvalue problem whose solution provides frequency and wavelength of eddies traveling over canopy.

@@

Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which there is no hydrodynamic instability (Fig(2)).
Variation of critical wave number with shear layer thickness for different grass height is shown in Figure \ref{K_vs_shear}

@@

@@

@@

Variation of critical wavenumber with shear layer thickness can be divided into three distinct region. First region, where critical wavelength does not increases linearly with the
increase of the shear layer thickness, this region occur for canopies with thick shear layer $H/\delta <10$. Second region named as shear layer region, where variation in critical
wavelength is proportional to the variation in shear layer thickness, which can be understood to be arising from a mechanism similar to Kelin-Helmholtz instability observed in the free
shear flow. Third region where critical wavelength does not vary with the variation of shear layer thickness. The boundary of these region depends on 

@@

The mechanism similar to Kelvin-Helmholtz instability is observed for a wide range of shear layer thickness $ <5H/\delta <800$ for canopy with low submergence ratio $h/H\leq 0.4$.
Canopy with low submergence ratio can also be understood to be close to the canopy in unbounded flow $h/H\sim0$ such as terrestrial canopy with atmospheric flow. 

@@

However the deviation of critical wavelength from the prediction based on of instability of shear layer in region-1 and region-3 for a given submergence ratio shows the existence of 

@@

comprising flow in shear layer and flow above the canopies. Whereas constant value of critical wavelength for all the shear layer smaller than a critical values shows the instability 

@@

by the collapse of eigen-modes for different shear layer thickness into a single profile as shown in Figure \ref{Asymptotic_mode}. The regime of shear instability of flow above the 

@@

Numerical investigation of flow in presence of canopy reveal three distinct zone of flow instability which might lead to monai. The occurrence of these three flow instability in a canopy

@@

\cite{Delangre06}\cite{Delangre04}\cite{Raupach96}\cite{Nepf99}

>>>>

transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87}\cite{Nepf99}. 
The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami \cite{AckermanOkubo93}, plays a central role
in the recruitment of microscopic marine organism such as blue mussel larvae \cite{Grizzle96}. The hydrodynamic mechanism underlying monami is the focus of this paper. 

@@

Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56}\cite{Raupach96}, have also been observed.

@@

Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 

@@

Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02}\cite{Raupach96} (Ghisalberti 2002, Raupach 1996) due to

@@

consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $, whereas observed critical 
Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires 

@@

In this paper we extend the idea of Raupach et.al\cite{Raupach96} via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grass

@@

instability associated with strong shear near the canopy top but also through shear instability of flow above grass canopies and shear instability of flow comprising flow above the canopy 
and flow in the shear layer. We further show that inclusion of drag due to grass predicts a threshold condition for waving. We confirm our analysis by comparing prediction from theory 
with field and experimental observations.

@@

primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depend weakly on flexibility of grass

@@

Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices. We use a mean field model for the coupling between the 
flow and the canopy. Drag force exerted by the grass is approximated via a 

@@

The steady uniform flow resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. Steady state flow driven by   

@@

\small$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$\normalsize arising from balance of drag force with pressure gradient with-in grass and a parabolic velocity 

@@

as a free shear layer in the previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 

@@

with drag along with matching shear at grass tip exerted by flow in unvegetated region, which leads to shear layer 
width \small $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity and
$H$ is half channel width.

@@

Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0$.

@@

R\left[\del_t+\del_x U \right]\grad^2\psi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2}{R}\frac{H^3}{\delta^3}U\psi_y\right)

@@

where $R= \frac{\rho U_0 H}{\mu}$ is Reynolds number. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, 
we get generalized Orr-Sommerfield equation 
which is an eigenvalue problem. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.

@@

@@

\label{Re_vs_delta}

@@

Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which the flow is stable 
(Figure\ref{Re_vs_delta}). The solution also predicts a critical wavenumber $k$ with maximum growth rate.
The variation of critical wave number with shear layer thickness for different grass height is shown in Figure \ref{K_vs_shear}. This variation of critical 
wavenumber with shear layer thickness can be divided into three distinct region. First, the region where critical wavelength does not increases linearly with the
increase of the shear layer thickness, this region occur for canopies with thick shear layer $H/\delta <10$. Second region identified as shear layer region, where variation in critical
wavelength is proportional to the variation in shear layer thickness, the instability in this region can be understood to arise from a mechanism similar to Kelin-Helmholtz instability 
observed in the free shear flow. Third, the region where critical wavelength does not vary with the variation of shear layer thickness. The boundary of these region depends on 

@@

\includegraphics[]{Low_drag_k}
\caption{Critical Wavenumber for different grass height as function of $f(\delta)$}
\label{Low_drag_k}

@@

The mechanism similar to Kelvin-Helmholtz instability is observed for a wide range of shear layer thickness $5 <H/\delta <800$ for canopy with low submergence ratio $h/H\leq 0.4$.
Canopy with low submergence ratio can also be related to the canopy in unbounded flow, such as terrestrial canopy with atmospheric flow with $h/H\sim 0$. 

@@

However, the deviation of critical wavelength from the prediction based on the instability of shear layer in region-1 and region-3 for a given submergence ratio shows the existence of 

@@

comprising flow in shear layer and flow above the canopies. The instability of flow comprising flow in shear layer and flow above the vegetation can be endorse by the linear 
variation of the critical wavelength with the $H-h+b\delta$ in Figure \ref{Low_drag_k}, where $b\delta$ is the amount of perturbation of the flow into the vegetation.
The constant value of critical wavelength for all the shear layer smaller than a critical value in region-3 suggest that the instability 

@@

by the collapse of eigen-modes for different $\delta<\delta_{cr}$ ($\delta/H=0.0037,0.0034,0.0024$ in Figure\ref{Asymptotic_mode}) into a single profile as shown in 
Figure \ref{Asymptotic_mode}, where $\delta_{cr}$ is the critical shear layer thickness
below which instability of vegetated flow arise through the mechanism of shear instability of flow in unvegetated region rather than through the mechanism of Kelvin-Helmholtz type 
instability of shear layer ($\delta/H=0.0052,0.0042$ in Figure \ref{Asymptotic_mode}). The regime of shear instability of flow above the 

@@

Numerical investigation of flow in presence of canopy reveal three distinct zone of flow instability which might lead to monami. The occurrence of these three flow instability in a canopy

@@

%\cite{Delangre06}\cite{Delangre04}\cite{Raupach96}\cite{Nepf99}
